SylvanSeeds, a database and web app to explore seed germination in temperate broadleaf and mixed forests

### 2020-06-21 doc/ms/v2_GEB/response-reviewers.Rmd

---
title: "Ref. GEB-2020-0155 - Response to reviewers"
author: "Eduardo Fernández Pascual"
date: "19/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please, find below my revision of the manuscript. I have considered and applied all suggestions. Considering the major issues, I have (1) expanded the explanation of the chosen biome and the creation of the species list; (2) clarified the missing full texts issue, providing a more detailed description, a PRISMA flow chart, and BibTeX files with the missing references; and (3) expanded the database to include the controlling factors required by the reviewers, such as substrate and number of containers.

# Response to the Editor

> I recommend that the manuscript be rejected with an opportunity to resubmit. The reviewers and myself agreed that the premise of the study was very interesting and relevant to the journal's audience, and appreciate the extensive work that went into its creation. However, a number of issues were identified, which I think preclude publication in present form, and whose resolution would potentially require extensive work on the part of the authors.

> First, the database's content and scope seems broader than warranted by its contents. The first reviewer and I both agree that the data do not necessarily provide full coverage of the proposed biome, and may contain species from beyond it. Additionally, I note that all of the sPlot records used to create a species list are from the Northern hemisphere, when the temperatre / broadleaf forest biome occurs widely in the southern hemisphere. While some of the identified studies end up sourcing data from southern hemisphere sites, it seems inaccurate and overly ambitious to suggest the dataset cover this biome. I am not sure that database with only northern hemisphere coverage of one biome, and with strong biases towards the species and sites represented in sPlot, would be of sufficient interest to receive support in the journal. I encourage the authors to expand the scope of their study, or alternatively provide a robust justification for the current scope, and a clearer title / narrative concomitant with the data.

The editor is right that the biome is also found in the southern hemisphere, although the extension there is smaller than in the northern hemisphere. In the new version of the manuscript, there is a longer explanation of why the NH biome was chosen, and how the species were selected. Summarising, the temperate broadleaf and mixed forests of the Northern Hemisphere were chosen as the subject of a sPlot project. Only the NH was chosen because it forms a coherent biogeographical unit, since it shared a common history during the Tertiary. The SH forests, on the other hand, are separated (and smaller) units and were not connected to the northern forests. This database was created for that specific sPlot project, and it was decided to do so because we expected that the number of germination references would be the highest, since many classical research groups have worked within the NH temperate broadleaf and mixed forests biome. This is developed in more detail in the manuscript and in the response to the reviewers.

In addition, the title has been changed to *Sylvan Seeds, a database and web app to explore the seed germination ecology of temperate broadleaf and mixed forests of the northern hemisphere*.

In any case, and in my personal view, I believe this is the more complete database of raw germination data for a coherent ecological unit and, spanning three continents of a densely populated and significant biome, it has enough global interest for the journal.

> Reviewer #1 also notes that 1/3 of the considered studies were omitted due to access issues; I suggest the authors try to resolve this point, or robustly clarify the justification for this choice. As the manuscript stands currently this text is not very convincing for a reader.

Please see the response to the reviewers about this point.

> Second, the first reviewer identified several issues related to the methodology used for the meta-analysis and literature search. I am sympathetic to these points and suggest the authors revisit their search criteria and data, or provide a robust response to these criticisms.

Please see the response to the reviewers about this point.

> As such, I do not think the data paper is ready to be accepted in present form, but am very willing to consider a thoroughly revised version.

# Responde to Referee 1

> The manuscript describes a dataset of species germination rates where the species were selected based on the sPlot species lists for temperate biomes. The shiny app is a nice addition that could be of use to the broader public with proper dissemination. It is interesting data to compile that will surely be valuable, but the scope of the biomes investigated and a lack of important methodological detail from the germination experiments limits the utility of the data. These concerns and a few others are detailed below.

> I wonder if defining the biome as “temperate broadleaf and mixed forests” is appropriate. There appears to be a number of species that generally occur in grasslands and biomes/habitats outside of forests (e.g. Dactylis glomerata, Achillea millifolium, Festuca ovina, etc.). Did each paper specify that the populations of seed collection were forested? If not, I recommend using something like a Whittaker biome diagram to describe the biomes as temperate rain, deciduous and dry forests, and grass/shrublands. Or just add additional biomes from the WWF classifications.

The new version of the manuscript contains a more detailed explanation of the reason to select the biome, and of the methodology used to construct the list of species. Please see the response to the editor and the manuscript for more comments on this.

The TBMF biome (in the northern hemisphere) is the area which was used to create the list of frequent species, for which the database was then compiled.  Biomes are defined as *relatively large units of land containing a distinct assemblage of natural communities and species, with boundaries that approximate the original extent of natural communities prior to major land-use change*. Thus, the TBMF biome does not contain only forests, but many other habitats such as heathlands, meadows, alpine grasslands, mires, dunes, etc. However, the plots used to create the list were filtered to be only forested plots (i.e., they were selected when they met the criterion that tree species dominated their plant cover). The list that has been produced includes species that occur with high frequency in forested plots of the TBMF biome in the nothern hemisphere. These are trees, shrubs, and herbaceous understory plants adapted to shade. But also species of forest margins, openings, riparian forests, ruderalized forests, etc.

The species mentioned by the reviewer are indeed species of the TBMF biome, and, even if today they may have their optimum in semi-natural meadows and pastures, they can occur also in forested plots (and in fact they do, as sown by the vegetation relevés we used). *Dactylis glomerata* and *Achillea millifolium* are species of mesic meadows (*Molinio-Arrhenatheretea*), which are secondary communities developped by human intervention over habitats that would be naturally occupied by temperate forests. The traditional interpretation of the origin of these communities is that they originated in forest margin communities that colonized the openings created by humans. Therefore, it is not surprising that these species are recorded with relative frequency in forest plots, as indeed they are. Similarly, it is not surprising that dry grassland (*Festuco-Brometea*) species such as *Festuca ovina* are also recorded in forest plots, especially in contact areas, as these are also communities that expanded over landscapes that would naturally be forests.

I agree that the seed lot coordinates that are provided with the database could be used to construct a Whittaker diagram of the seed lots, or alternatively to asign them to WWF biomes. But I feel that this is more of an analytical usage of the data, which can be done by any user with the information provided in the dataset. Still, I am willing to do such tasks if the editor finds them essential.

> The justification for restricting the search to certain biomes and species is weak (Lines 49-50), especially given sPlot is a global database. If picking these biomes covers most of the available data, why not extend beyond this biome given the added work is relatively minimal? Then the dataset would be truly global. There are a lot of papers on germination outside of this biome that could make for valuable comparisons (e.g., tundra/alpine biomes from Vandvik et al. 2017. Botany 95:73-88; Tropical forests from Yu et al. 2008. Plant Ecology 297:1-16). Why not expand the search to all species within sPlot? If the search is expanded to all species in sPlot, does the number of articles requiring processing become unreasonable? This would make for a better justification for restricting the search.

I have included a more detailed description of the reasons behind chosing the TBMF biome to construct the database: *The building of the dataset was part of a research project conducted within the sPlot initiative (Bruelheide et al., 2019), specifically sPlot’s project #12 (https://www.idiv.de/en/sdiv/working_groups/wg_pool/splot/projects.html). The goal of sPlot #12 was to study taxonomical and functional plant diversity in deciduous forests, which are dominant in temperate latitudes with mild to cold winters and relatively warm and wet summers. In the northern hemisphere, they also form a coherent evolutionary and biogeographic unit, sharing a common Tertiary history until they became separated between North America, Asia and Europe (Milne & Abbott, 2002).*

Completing the database for all the species in sPlot is a worthwhile long-term goal (as mentioned in the *Utility of the database* section of the manuscript), and one that I am interested in contributing to in the future. But it was completely beyond the resources I had to construct this database, which depended on a 2-year personal grant. It took me 9 months of focused work to build a database for 1,393 species. There are 54,519 species in sPlot 2.1 (*Bruelheide, H., Dengler, J., Jiménez‐Alfaro, B., Purschke, O., Hennekens, S. M., Chytrý, M., ... & Aubin, I. (2019). sPlot–A new tool for global vegetation analyses. Journal of Vegetation Science, 30(2), 161-186*). Therefore, a rough estimate is that it would take me 29 years of work to do a database for all sPlot species! Even if the time would probably be shorter (because I do expect the number of references would be lower outside the TBMF biome), the number of references would be expanding each year with new research. So, because one has to start somewhere, I chose to do it in connection to sPlot project 12, which focused on temperate broadleaf and mixed forests of the northern hemisphere; and I chose to do it because this biome has received much attention in seed research.

> The literature search would benefit from more description and/or investigation. What made a full text impossible to access? Not having access to a few full texts is understandable, but 236 would make up more than 1/3 of the 611 that were to be relevant. This data could be very important to the completeness of the dataset. Using PRISMA guidelines for proper meta-analysis data gather is highly recommended (http://www.prisma-statement.org).

Just to be precise, the proportion of not-found information is actually smaller. Of 1489 references for which I searched for the full text, I found 1253 (NOTE: While doing the revisions, I have found and corrected small discrepancies in the numbers of references, caused because a few of the records were removed by a mismatch during taxon name standardization). Of these, 639 (51%) were not relevant. If we assume the same proportion of relevant references in the cases for which I could not find the full text, we can estimate that there were 116 relevant articles which I could not include in the database: 16% of the potentially relevant references (614 found + 116 estimated not found = 730; 116/730 =  0.16). Furthermore, I must note that information is not homogenesouly distributed thorough the references: there are many references for some popular species, and fewer references for rarer species. Therefore the amount of valuable information does not increase linearly with increasing number of references, and it is very likely that a large proportion of that 16% is made up of references considering species which are already in the database. In conclusion, I believe that the amount of missing information is within reasonable limits.

Be that as it may, I have improved the description of data gathering in the text, including a description of the process of finding full texts: *An effort was made to access the full text of all these references, using in order the following channels: (i) “Find Full Text” functionality in EndNote X9, (ii) online access to the journals using the institutional subscriptions of the University of Oviedo; (iiI) request to the authors through ResearchGate; (iv) general search of pdf files in Google using the article details. Unfortunately, 236 references could not be accessed. Most of these belonged to smaller publishers not included in the subscriptions of the University of Oviedo, and to older journal volumes which had not been digitalised.*

I have also included some supplementary materials to describe the data gathering: the boolean search string I used in the Web of Science, and a PRISMA flow chart.

Additionally, I have added the lists of the three types of references (non-relevant, found and not-found) as BibTeX supplementary files. In this way, it will be possible to check what I did, and to complete it in the future by accessing the references that I could not find.

> There appears to be a lot of important information missing in the dataset that would allow a user of the dataset to control for methodological artifacts or biases. For example: what kind of substrate were the seeds germinated in (Agar? Soil? Sand?)? What was the density of seeds sown? How large were the containers holding the seeds? How many replicates were there to determine the number germinated versus germinable? Were any treatments, such as drought or nutrient addition, used in the experiment? Were the seeds sown in a monoculture?

As requested, I have revised the references to retrieve the requested information, and added it to the database: type of substrate; type and dimensions of the containers; number of containers and number of seeds sown per container; treatments of sterilization and nutrients. Density of seeds is not reported in the references, but it can be estimated from the dimension of the containers and the number of seeds sown. All cases included in the database correspond to monoculture experiments, and special treatments (drought, heavy meatals, smoke, etc.) were not included; this is now stated in the text. Please see the manuscript for the description of the new variables in the database.

> Most of the studies would have replicated germination trials (e.g. sown 100 seeds across 5 replicates: 20 seeds per replicate). Therefore, there is likely true error estimates (and mean percent germination) that can be obtained from the papers. These are a highly desirable metrics from the papers and would improve the error estimates in the shiny app as well.

As requested, I have included the number of containers ("replicates") and number of seeds sown per replicate in the database.

> Line 12- add an “and” so that it reads “darkness, and constant or alternating”

Done.

> Line 16- Will the data collection be continuous? Maybe assign a final year here.

Done.

> Line 31- add an of; “all of these”

Done.

> Line 35- use commas instead of –

Done.

> Line 39-40- “which implies a set of derived complications in the analytical treatment of the data”- unclear what this means and how it applies to germination as a number of successes given seeds sown (trials).

I have removed the whole sentence.

> Line 45- Rather than “at this and that temperature”, try “across a temperature gradient”?

Done.

> Line 95- give a brief explanation of GA3

Done.

> Line 104-105- Remove highlight of Fig. 1.

Done.

> Line 108- incorrect degree symbol.

It was the degree symbol rendered in the Calibri font, the new manuscript is in Arial.

> Line 154- remove the “a” to read “can have wide applicability”

Done.

> Line 155-157- Try “The data can extend current trends in global analyses of plant traits at species (Diaz et al., 2016) and community levels (Bruelheide t al., 2018) to seed germination.”

Done.

> Line 160- Remove “to”

Done.

> Line 262- As the authors note, the references will need cleaning.

Indeed.

> Can the species names be searchable in addition to the drop-down menu in the shiny app?

The species name can be deleted, and then one can write in the box. This will trigger an autocomplete, which will allow to search.

> There are special character errors in the “Reference” column of the dataset: e.g., «ali?kan, O., et al. (2012). "Influences of presowing treatments on the germination and emergence of fig seeds (Ficus carica L.)." Acta Scientiarum. Agronomy 34(3): 293-297.

This should be fixed now.

# Responde to Referee 2

> This data paper introduces a new resource available to ecologists: a data base and associated app summarising seed germination data for a large number of plant species in the temperate broadleaf/mixed forest biome. The authors used an extensive literature search to identify sources of germination data, and have gone through a substantial effort in extracting data from a large number of published articles. The resulting ‘Sylvan Seeds’ data base and app has the potential to be a valuable tool for researchers and applied foresters. I had a quick play around with the app – it provides nice species-specific data summaries (although I think the data base itself is going to be a more useful resource).

> I have little to criticise in terms of the approaches used (which seem sound), but believe that the presentation would benefit from a few changes to make it more reader friendly and accessible for people whose main interest might not be in seed ecology and to potential non-academic users of the database/app:

> L1 (Title): I am not sure how many people will immediately know what a ‘shiny app’ is. (I did not...) Could another more general term be used?

I changed it to *web app*.

> L5 (Motivation): This section could perhaps start with a sentence on why seed germination is an important phenomenon in ecology?

Done.

> L9-10: “Data has been gathered using a meta-analytical approach to search the literature.” Meta-analytical approach may suggest that some kind of data analysis was involved. Could be changed to something along the lines of “Data were gathered using a systematic literature search approach” or similar. (Later in the text you could then mention that the exact approach taken was inspired by approaches typically used in meta-analyses?)

Changed as suggested.

> L14: “across the biome and beyond” – unclear meaning

Changed to *Seed lots collected throughout the biome and in neighbouring areas*.

> L16: today -> March 2019?

Changed to the year of the most recent seed lot, 2017.

> L17: frequent -> common?

I believe it is more appropiate to use frequent, as frequency is the common term in plant ecology: *Frequency is the number of times a plant species occurs in a given number of quadrats*. Frequency, thus defined, was the criterion used to include species in the search.

> L20: wide -> wider?

Changed.

> L25: Again, considering the broad readership of Global Ecology and Biogeography, it might be helpful to start with a sentence or two on the importance of studying seed traits/germination/seed ecology (and – given the journal – ideally also something on why a global approach can contribute to the field of seed ecology).

Done.

> L41: scarcity of -> lack of? (Alternatively, if such data bases exist, but there are only few of them, stick to original wording but mention which ones exist.)

Changed to *lack of*.

> L45: at this and that temperature -> at various temperatures

Changed to *accross a temperature gradient* as suggested by referee 1.

> L49-50: home of many of the classical research groups in seed ecology -> research focus of (…) or similar?

Changed to *At the same time, a relatively high number of classic seed biology research groups have worked within the TBMFNH areas, e.g. in the USA (Baskin, 1995), the UK (Grime et al., 1981), Belgium (Van Assche & Vandelook, 2006) and Japan (Washitani, 1988). *.

> L51: see previous comment regarding reference to meta-analysis

Changed to *To gather data for the project, a systematic literature search (Koricheva et al., 2013) was applied to a list of frequent species representing the flora of the TBMFNH.*

> L57: not all readers might be familiar with the term vegetation relevé -> provide brief explanation?

The following explanation has been added: *(i.e. records of plant species co-occurring in a sampling plot)*

> L58: the reference to sPlot’s project #12 might not mean much to people not involved with sPlot. -> Provide link to website or further explanation?

I have included the following description: *The building of the dataset was part of a research project conducted within the sPlot initiative (Bruelheide et al., 2019), specifically sPlot’s project #12 (https://www.idiv.de/en/sdiv/working_groups/wg_pool/splot/projects.html). The goal of sPlot #12 was to study taxonomical and functional plant diversity in deciduous forests, which are dominant in temperate latitudes with mild to cold winters and relatively warm and wet summers. In the northern hemisphere, they also form a coherent evolutionary and biogeographic unit, sharing a common Tertiary history until they became separated between North America, Asia and Europe (Milne & Abbott, 2002).*. Unfortunately, no articles derived from sPlot #12 have been finished so far, so I cannot cite them, which would be the ideal solution.

> L72: ‘To obtain a final list of species (…)’. The primary reason for focusing on the subset of species that were present in a minimum of 5% of the relevés of an ecoregion was presumably to make the work achievable given the large number of species (>7000) in the full species pool?

There was the practical reason mentioned by the reviewer, but also a methodological consideration. I have added the following explanation: *There were two reasons to use this frequency threshold: (i) it made the search achievable with the resources available; and (ii) it filtered out rare species that may not be representative of TBMFNH, and which could have been recorded for a variety of reasons (wrong taxonomical identifications, mistakes in plot coordinates, border areas, unnatural areas such as parks, etc.).*

> L81-82: “(…) for 236 it was impossible to access the full text.” Unclear meaning: Was this because of lack of institutional licenses for accessing journals, or because of something else?

Please see the response to referee 1 about the same question. The causes are now described as follows: *An effort was made to access the full text of all these references, using in order the following channels: (i) “Find Full Text” functionality in EndNote X9, (ii) online access to the journals using the institutional subscriptions of the University of Oviedo; (iiI) request to the authors through ResearchGate; (iv) general search of pdf files in Google using the article details. Unfortunately, 236 references could not be accessed. Most of these belonged to smaller publishers not included in the subscriptions of the University of Oviedo, and to older journal volumes which had not been digitalised.*.

### 2020-06-21 doc/ms/v2_GEB/response-reviewers.Rmd

> L95: GA3 -> gibberellic acid (to help readers who are less familiar with the topic, could perhaps also mention that GA3 speeds up seed germination)

Done.

> L96: the term ‘response variable’ suggests that you conducted some kind of statistical analysis. -> Delete term or change wording.

Removed *the response variable* part of the sentence.

### 2020-06-21 doc/ms/v3_NewPhytol/ms.Rmd
NP submission

---
title: "Sylvan Seeds, a database and web app to explore the seed germination ecology of temperate broadleaf and mixed forests of the northern hemisphere"
subtitle: "Running title: Sylvan Seeds, a forest germination database"
output: github_document
bibliography: refs.bib
csl: global-ecology-and-biogeography.csl
link-citations: yes
always_allow_html: true
---

@@

# Abstract

*Motivation* Seed traits have functional significance in all levels of plant ecology, from population dynamics to community assembly and species distribution models. Recent discussions on the utility of seed traits in ecology have highlighted the unavailability of reliable germination databases of wide geographical scope. This data paper presents a first global dataset of raw germination data, encompassing an ecologically and biogeographically coherent unit: the Temperate Broadleaf and Mixed Forests biome of the northern hemisphere. Data were gathered using a systematic literature search approach.

*Main types of variable contained* Proportion of seeds germinated in laboratory experiments, under different experimental combinations of scarification, stratification, light or darkness, and constant or alternating germination temperatures.

*Spatial location and grain* Seed lots collected throughout the biome and in neighbouring areas, with approximate geographical coordinates of the collecting sites.

*Time period and grain* Seed lots collected between 1920 and 2017.

*Major taxa and level of measurement* 326 plant species with high frequency in vegetation relevés of the biome, representing 74 families of seed plants (gymnosperms and angiosperms).

*Software format* The database is provided as a single *.csv* file. The shiny web app *Sylvan Seeds* is also presented, to explore the database and make it accessible to the wider public.

# Keywords

Temperate deciduous forests, temperate evergreen forests, temperate coniferous forests, seed traits, germination database, seed dormancy, germination temperature, alternating temperatures, light germination, dark germination

@@

Seed traits have functional significance in all levels of plant ecology, from population dynamics [@RN3043] to community assembly [@RN2942] and species distributions [@RN2942]. Recent discussions [@RN2352; @RN4655] highlighted the unavailability of reliable germination databases of wide geographical scope as a major limitation to produce a global synthesis of the seed ecological spectrum. Seed germination is the first major developmental transition in the life of a plant, and as such it has profound effects on the plant's ecology [@RN2384]. As such, the timing of this transition is controlled by a combination of environmental signals. Amongst these are temperature [@RN2249], diurnal temperature alternation [@RN2217], light [@RN4656] and seed dormancy inductors and relievers [@RN3063]. All of these signals interact to produce a coarse- and fine-scale regulation of germination timing, integrating inputs from both seasonal climatic cycles [@RN3132] and local environmental gradients [@RN2865]. The practical consequence of this is that the response of seeds to, for example, light will depend on other conditions set by the experimenter. This makes it difficult to summarise germination “traits” into a single value in a way comparable to seed mass, specific leaf area or plant height [@RN2923], and may explain the lack of global germination databases. Germination compilations with a biogeographical background, of which the prime example today is the cornerstone book of Baskin and Baskin [-@RN3214], only provide summary information, for instance the interpreted optimal germination temperature instead of the proportion of seeds germinated across a temperature gradient.

To fill this gap, this data paper presents a first global dataset of raw germination data for an ecologically and biogeographically coherent unit: the Temperate Broadleaf and Mixed Forests biome of the northern hemisphere (*TBMFNH*). The building of the dataset was part of a research project conducted within the *sPlot* initiative [@RN4658], specifically *sPlot*’s project #12 (https://www.idiv.de/en/sdiv/working_groups/wg_pool/splot/projects.html). The goal of *sPlot* #12 was to study taxonomical and functional plant diversity in deciduous forests, which are dominant in temperate latitudes with mild to cold winters and relatively warm and wet summers. In the northern hemisphere, they also form a coherent evolutionary and biogeographic unit, sharing a common Tertiary history until they became separated between North America, Asia and Europe [@RN4684]. At the same time, a relatively high number of classic seed biology research groups have worked within the *TBMFNH* areas, e.g. in the USA [@RN1045], the UK [@RN3273], Belgium [@RN656] and Japan [@RN1188]. For this reason, *sPlot* project #12 was chosen to incorporate a biome-wide analysis of seed traits, as it was expected that the *TBMFNH* would provide a wide scope of available data. To gather data for the project, a systematic literature search [@RN2401] was applied to a list of frequent species representing the flora of the *TBMFNH*. This data paper makes the database accessible to the research community. Additionally, the *Sylvan Seeds* shiny web app is presented to facilitate the exploration of the data.

# Methods

@@

The list of frequent species (plus the synonyms recorded in the relevés) was incorporated into a Boolean search string, together with the words "(seed OR seeds) AND (dormancy OR germination)” (**Supplementary material 1**). This string was searched in the *Thompson Reuters Web of Science* on 5 Mar 2019, returning 6,791 results. The processing of these results is described in a PRISMA flow diagram (**Supplementary material 2**). A first filter of the results by the relevance of the title retained 1,489 references. An effort was made to access the full text of all these references, using in order the following channels: (i) “Find Full Text” functionality in *EndNote X9*, (ii) online access to the journals using the institutional subscriptions of the University of Oviedo; (iiI) request to the authors through *ResearchGate*; (iv) general search of *pdf* files in *Google* using the article details. Unfortunately, 236 references could not be accessed. Most of these belonged to smaller publishers not included in the subscriptions of the University of Oviedo, and to older journal volumes which had not been digitalised. The resources of this project did not make it possible to further pursue these references, but they are included as a *BibTeX* file (**Supplementary material 3**) in case they can be processed in the future. The found texts were accessed to retrieve data for the database. At this stage, references were considered to be non-relevant if they (i) did not pertain to species in the original frequent species list; (ii) did not study germination in species monocultures (e.g. competition studies); (iii) did not report the minimal information needed (number of seeds sown, germination percentage and germination temperature). Records were also excluded when they pertained to special germination treatments that do not allow for general comparisons (i.e. use of germination promoters such as the plant hormone gibberellic acid; drought treatments, including polyethylene glycol solutions; application of heavy metals or other pollutants; smoke-related germination promoters such as karrikins). Of the accessed references, 639 were non-relevant (*BiBTeX* file in **Supplementary material 4**) and 614 provided relevant data (*BiBTeX* file in **Supplementary material 5**).

@@

# Description of the database

@@

The final database contains `r MSrecords` records (germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) from 614 references. The plant materials had been collected across the Temperate Broadleaf and Mixed Forest biome and surrounding areas (**Fig. \@ref(fig:fig1)**) in both hemispheres. The oldest record was from `r MSoldest` and the top three contributing countries were the USA (`r MScountries[1, 2]`), the UK (`r MScountries[2, 2]`) and Japan (`r MScountries[3, 2]`). There were `r MSspecies` species represented, from `r MSfamilies` seed plant families. The total estimate of seeds used in the experiments was `r MSseeds`. The range of germination temperatures (weighted average of the daily thermoperiod) went from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures and `r MSaltN` of alternating temperatures. Light was used in `r MSlightY` records, darkness in `r MSlightN` and `r MSlightNA` did not provide information on this parameter. The experiments were performed with unstratified seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification. Scarification was applied to `r MSscarY` records.

@@

# Data accessibility statement

All persons can use the database providing they cite this data paper properly in any publications or in the metadata of any derived products that are produced using the database. The database is provided as supplementary material and will be stored in *Dryad*. It can be visualized with the *Sylvan Seeds* app at http://sylvanseeds.shinyapps.io/sylvanseeds/. The code of the app is stored at (note: the GitHub page is kept private until publication of the manuscript).

@@

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Example of the germination records as shown by the *Sylvan Seeds* app. Records for one species, the European pedunculate oak, *Quercus robur*. Each panel shows the results for a combination of experimental conditions, with the germination temperature varying within each panel. Bars represent the mean germination proportion and brackets the 95% binomial confidence interval.", fig.align = "center", fig.width=7, fig.height=9}
p2
```

>>>>

---
title: "SylvanSeeds, a database and web app to explore seed germination in temperate broadleaf and mixed forests"
subtitle: "Running title: SylvanSeeds, a forest germination database"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
---

@@

**Author:** Eduardo Fernández-Pascual **Institutional address:** Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, C/ Catedrático Rodrigo Uría, 33006 Oviedo/Uviéu, Spain **Email:** efernandezpascual@gmail.com **Telephone:** +34985104787

**Total word count** 2,578 **Introduction** 530 **Materials and Methods** 968 **Results** 848 **Discussion** 226 **Number of figures** 2 in colour **Tables** 0 **Supporting information** 6

# Summary

* Seed traits have functional significance in all levels of plant biology, but there is a lack of germination databases of wide geographical scope. This article presents *SylvanSeeds*, a first global dataset of germination records for an ecologically coherent unit: Temperate Broadleaf and Mixed Forests.

* Data were gathered with a systematic literature search. A list of frequent taxa of the study area was created using `r MSplots` vegetation relevés from the sPlot database. The list was searched in the Web of Science. 6,791 references were screened, finding 614 articles from which data were extracted.

* *SylvanSeeds* includes `r MSrecords` germination records of `r MSspecies` species from `r MSfamilies` families (gymnosperms and angiosperms), collected in 46 countries between 1920-2017. It provides raw data for meta-analysis: proportions of seeds germinated in laboratory experiments of scarification, stratification, light-darkness, and constant-alternating temperatures. *SylvanSeeds* is freely distributed as a *.csv*. The shiny app *SylvanSeeds* is also presented, to make data accessible to the public.

* *SylvanSeeds* advances functional seed ecology and brings two innovations to plant science.  First, the data-gathering methodology can be extended to other biomes. Second, database and app can be a standard in further efforts to compile germination data.

**Keywords:** germination database, germination temperature, light and dark germination, seed dormancy, seed traits, temperate coniferous forests, temperate deciduous forests, temperate evergreen forests

@@

Seed traits have functional significance in all levels of plant biology, from population dynamics [@RN3043] to community assembly [@RN2942] and species distributions [@RN2942]. Recent discussions [@RN2352; @RN4655] highlighted the lack of reliable germination databases of wide geographical scope as a major limitation to produce a global synthesis of the seed ecological spectrum. Seed germination is the first major developmental transition in the life of a plant, and it has profound effects on the plant's ecology [@RN2384]. As such, the timing of this transition is controlled by a combination of environmental signals. Amongst these are temperature [@RN2249], diurnal temperature alternation [@RN2217], light [@RN4656] and seed dormancy inductors and relievers [@RN3063]. All of these signals interact to produce a coarse- and fine-scale regulation of germination timing, integrating inputs from both seasonal climatic cycles [@RN3132] and local environmental gradients [@RN2865]. The practical consequence of this is that the response of seeds to, for example, light will depend on other conditions set by the experimenter. This makes it difficult to summarise germination “traits” into a single value in a way comparable to seed mass, specific leaf area or plant height [@RN2923], and may explain the lack of global germination databases. Germination compilations with a biogeographical background, of which the prime example today is the cornerstone book of Baskin and Baskin [-@RN3214], only provide summary information, for instance the interpreted optimal germination temperature instead of the proportion of seeds germinated across a temperature gradient.

To fill this gap, this article presents *SylvanSeeds*, a first global dataset of raw germination data for an ecologically and biogeographically coherent unit: the Temperate Broadleaf and Mixed Forests biome of the northern hemisphere (*TBMFNH*). The building of the dataset was part of a research project conducted within the *sPlot* initiative [@RN4658], specifically *sPlot*’s project #12 (https://www.idiv.de/en/sdiv/working_groups/wg_pool/splot/projects.html). The goal of *sPlot* #12 was to study taxonomical and functional plant diversity in deciduous forests, which are dominant in temperate latitudes with mild to cold winters and relatively warm and wet summers. In the northern hemisphere, they also form a coherent evolutionary and biogeographic unit, sharing a common Tertiary history until they became separated between North America, Asia and Europe [@RN4684]. At the same time, a relatively high number of classic seed biology research groups have worked within the *TBMFNH* areas, e.g. in the USA [@RN1045], the UK [@RN3273], Belgium [@RN656] and Japan [@RN1188]. For this reason, *sPlot* project #12 was chosen to incorporate a biome-wide analysis of seed traits, as it was expected that the *TBMFNH* would provide a wide scope of available data. To gather data for the project, a systematic literature search [@RN2401] was applied to a list of frequent species representing the flora of the *TBMFNH*. This data paper makes the database accessible to the research community. Additionally, the *SylvanSeeds* shiny web app is presented to facilitate the exploration of the data.

# Materials and Methods

@@

The list of frequent species (plus the synonyms recorded in the relevés) was incorporated into a Boolean search string, together with the words "(seed OR seeds) AND (dormancy OR germination)” (**Supporting Information 1**). This string was searched in the *Thompson Reuters Web of Science* on 5 Mar 2019, returning 6,791 results. The processing of these results is described in a PRISMA flow diagram (**Supporting Information 2**). A first filter of the results by the relevance of the title retained 1,489 references. An effort was made to access the full text of all these references, using in order the following channels: (i) “Find Full Text” functionality in *EndNote X9*, (ii) online access to the journals using the institutional subscriptions of the University of Oviedo; (iii) request to the authors through *ResearchGate*; (iv) general search of *pdf* files in *Google* using the article details. Unfortunately, 236 references could not be accessed. Most of these belonged to smaller publishers not included in the subscriptions of the University of Oviedo, and to older journal volumes which had not been digitalised. The resources of this project did not make it possible to further pursue these references, but they are included as a *BibTeX* file (**Supporting Information 3**) in case they can be processed in the future. The found texts were accessed to retrieve data for the database. At this stage, references were considered to be non-relevant if they (i) did not pertain to species in the original frequent species list; (ii) did not study germination in species monocultures (e.g. competition studies); (iii) did not report the minimal information needed (number of seeds sown, germination percentage and germination temperature). Records were also excluded when they pertained to special germination treatments that do not allow for general comparisons (i.e. use of germination promoters such as the plant hormone gibberellic acid; drought treatments, including polyethylene glycol solutions; application of heavy metals or other pollutants; smoke-related germination promoters such as karrikins). Of the accessed references, 639 were non-relevant (*BiBTeX* file in **Supporting Information 4**) and 614 provided relevant data (*BiBTeX* file in **Supporting Information 5**).

@@

# Results

@@

The final database contains `r MSrecords` records (germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) from 614 references. The plant materials had been collected across the Temperate Broadleaf and Mixed Forest biome and surrounding areas (**Fig. \@ref(fig:fig1)**) in both hemispheres. The oldest record was from 1920 and the top three contributing countries were the USA (`r MScountries[1, 2]`), the UK (`r MScountries[2, 2]`) and Japan (`r MScountries[3, 2]`). There were `r MSspecies` species represented, from `r MSfamilies` seed plant families. The total estimate of seeds used in the experiments was `r MSseeds`. The range of germination temperatures (weighted average of the daily thermoperiod) went from `r MSminT` to `r MSmaxT` ºC, with `r MSaltY` records of constant temperatures and `r MSaltN` of alternating temperatures. Light was used in `r MSlightY` records, darkness in `r MSlightN` and `r MSlightNA` did not provide information on this parameter. The experiments were performed with unstratified seeds in `r MSstratN` records, and of the rest, the majority (`r MSstratCold`) went through cold stratification. Scarification was applied to `r MSscarY` records.

@@

# Supporting Information

**Supporting Information 1** Boolean search string used in the Web of Science. **Supporting Information 2** PRISMA 2009 flow chart describing the process of reference screening and inclusion. **Supporting Information 3** BibTeX bibliography listing the not found full texts.  **Supporting Information 4** BibTeX bibliography listing the non-relevant references.  **Supporting Information 5** BibTeX bibliography listing relevant references included in the *SylvanSeeds* database.  **Supporting Information 6** *SylvanSeeds* database in a single *.csv* file.

# Data accessibility

All persons can use the database providing they cite this paper properly in any publications or in the metadata of any derived products that are produced using the database. The database is provided as Supporting Information. It can be visualized with the *SylvanSeeds* app at http://sylvanseeds.shinyapps.io/sylvanseeds/. The code of the app is stored at (note: the GitHub repo is kept private until publication of the manuscript).

@@

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Example of the germination records as shown by the *SylvanSeeds* app. Records for one species, the European pedunculate oak, *Quercus robur*. Each panel shows the results for a combination of experimental conditions, with the germination temperature varying within each panel. Bars represent the mean germination proportion and brackets the 95% binomial confidence interval.", fig.align = "center", fig.width=7, fig.height=9}
p2
```
